### 2021-03-12 documentation/dear_diary/log_diary.tex

\documentclass[]{article}
\usepackage[utf8]{inputenc}
\usepackage[danish,english]{babel}
\usepackage{graphicx}
\usepackage[usenames]{xcolor}
\usepackage{times}
\usepackage{listings}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{textpos}
\usepackage{pgfgantt}
\usepackage{lscape}
\usepackage{indentfirst}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{ifthen}
\usepackage{float}
\usepackage{pict2e}
\usepackage[export]{adjustbox}
\pgfplotsset{compat=1.13}
\usepackage{siunitx}
\usepackage{tabularx}

% adjustment to page format
\marginparwidth=0pt
\oddsidemargin=0pt
%\evensidemargin=0pt
\marginparsep=0pt
%\topmargin=-15 mm
\textwidth=168 mm
\textheight=210 mm
\parskip 0.27 em
\parindent=0pt
%\leftmargin=1cm
\pagenumbering{arabic}
% Define DTU color
\definecolor{dtured}{RGB}{153,0,0}
\title{Projektplan}

\begin{document}
This log shall be used as a technical diary we write throughout the course. It will be used to write the final report more easily.

\section{05-03}
We've set up PlatformIO in Visual Studio Code as the environment in which we write our code.

\section{10-03}
We've made an adjustment in the ressources and will be using a teensyboard 3.6 instead of a 4.1, since the 3.6 board is already in stock at DTU. It has a slower clock (180Mhz, compared to 4.1's 600Mhz), but we judge that this won't matter to the project solution.

\end{document}